---
title: "p8105_hw3_ss7636"
output: html_document
date: "2025-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
library(p8105.datasets)
data("instacart")
instacart
```
The dataset contains 1,384,617 observations of 131,209 unique users, where each row in the dataset is a product from an order. There is a single order per user in this dataset. There are 15 variables in this dataset, and a couple key important variables are order_id: order identifier, product_id: product identifier, and add_to_cart_order: order in which each product was added to cart

```{r}
n_aisles = instacart |> 
  summarize(n_aisles = n_distinct(aisle)) |> 
  pull(n_aisles)

n_aisles

```

```{r}
intacart_items_ordered = instacart |> 
  count(aisle, name = "items_ordered") |> 
  arrange(desc(items_ordered))

intacart_items_ordered
```

There are 134 aisles and most items are ordered from the fresh vegetables and fresh fruits sections. 

```{r}
intacart_items_ordered  |> 
  filter(items_ordered > 10000) |> 
  ggplot(aes(x = reorder(aisle, -items_ordered), y = items_ordered)) + 
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # rotate labels
  labs(title = "Items Ordered per Aisle (Above 10,000 Orders)",
       x = "Aisle",
       y = "Number of Items Ordered")

intacart_items_ordered
  
```

```{r}
top_3_items = instacart |> 
  filter(aisle == c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  select(aisle, product_name) |> 
  group_by(aisle, product_name) |> 
  summarize(n_orders = n()) |> 
  arrange(aisle, desc(n_orders)) |> 
  filter(row_number() <= 3) 

top_3_items
```

```{r}
knitr::kable(top_3_items)
```


```{r}
mean_hour = instacart |> 
  filter(product_name == c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  select(product_name, order_hour_of_day, order_dow) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = round(mean(order_hour_of_day), 2)) |> 
  pivot_wider (
    names_from = "order_dow",
    values_from = "mean_hour"
  )

mean_hour
```
```{r}
knitr::kable(mean_hour)
```



